DRAFT: Harmonisation and trait aggregation paper

### 2020-12-18 Paper/Draft_harmonization_paper.tex
18.12: Abstract draft and figure correction

@@

In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

@@

\begin{sidewaysfigure}
  \label{fig:boxplots_scores_on_constrained_axis}
  \centering
  \includegraphics[width= 22cm, height=12cm]{Scores_on_constrained_axis_combined.pdf}
  \caption{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled.}
\end{sidewaysfigure}

@@

With the increasing use of computer vision in ecology, both species identification and trait quantification could be automated \cite{milosevic_application_2020, hoye_deep_2020}. This might result in more observational data and trait data on species-level in the future. However, if traits need to be aggregated we showed that trait aggregation is often in conclusion with expert assignments, specifically aggregation approaches using the median. Which aggregation method the objectively better measure is cannot be answered, but when trait aggregation is necessary 1) median aggregation seems to be favorable, 2) highly variable traits and taxa where a few genera have a high number of species should be treated carefully, 3) coding of traits should be considered, and 4) weighting trait information based on the taxa present in the used trait database does not seem to have much influence. No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits, a measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate.

>>>>

\section*{Summary}

% FROM FRESHWATER BIOLOGY Website:
% All papers should include a summary, in short numbered paragraphs, limited to about 3% of the length of the text, and in any case to not more than 500 words. This should provide a concise statement of the scope of the work and its principal findings and be fully intelligible without reference to the main text. It is helpful to structure your abstract as follows: point 1 = background to and reason for the study and the aims; point 2 = brief description of methods used; point 3 = main project results (address the aims); point 4 = conclusions of your study; point 5 = the contribution that your results make to the field / why your study is useful to an international audience. Referring to abstracts in recently published issues may help with structuring an abstract.

1. Invertebrate traits are increasingly used in ecological studies because they provide a mechanistic understanding of species-environment relationships and are suitable for large scale analysis due to their consistent response to environmental gradients. The development of invertebrate trait databases for different continents in recent years should have created impetus for trait-based research, especially on large scales. However, a number of data-related problems hamper the work with trait-based approaches. For example, standardised trait definitions are lacking and coding of traits varies across trait databases impeding trait data synthesis. Another issue is different taxonomic resolution between observational taxonomic data and trait databases, leading to the need to aggregate traits. 
% Aim of our study?

2. In this study, we describe discrepancies in trait definitions of trait databases from Europe, North America, New Zealand and Australia for traits frequently used in ecological studies. Based on these databases we established 4 novel trait datasets by harmonising 7 commonly used grouping features. Two of these datasets were used to evaluate different trait aggregation methods by comparing aggregated traits with assigned traits by experts at family-level. Furthermore, we explored the effect of grouping feature harmonisation and trait aggregation by re-analysing a study on the effect of anthropogenic salinisation on biological traits using harmonised grouping features and aggregated traits.

3. We found that investigated trait databases are distinct with regard to how grouping features are differentiated into traits, the scope of the grouping feature definitions, how traits are defined, and the coding to describe traits. Comparison of aggregated and assigned traits showed that aggregated traits are generally in conclusion with expert assignments, especially approaches using the median. Grouping feature harmonisation and trait aggregation had a minor impact in the re-analysis, although averaging effects can occur due to harmonisation diminishing the signal of a trait.

4. By outlining discrepancies in trait definitions we hope to motivate the development of a standardized terminology for invertebrate traits. When trait aggregation is needed, our results provide guidance under which circumstances the choice of a trait aggregation method becomes important.

@@

In this study, we follow the terminology proposed by Schmera et al. \cite{schmera_proposed_2015}, where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

@@

% \begin{sidewaysfigure}
%   \label{fig:boxplots_scores_on_constrained_axis}
%   \centering
%   \includegraphics[width= 22cm, height=12cm]{Scores_on_constrained_axis_combined.pdf}
%   \caption{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled.}
% \end{sidewaysfigure}

\begin{figure}[H]
  \label{}
  \centering
  \includegraphics[width=16.5cm, height=25cm]{Scores_on_constrained_axis_comb_ALT.pdf}
  \caption{}
\end{figure}

@@

With the increasing use of computer vision in ecology, both species identification and trait quantification could be automated \cite{milosevic_application_2020, hoye_deep_2020}. This might result in more observational data and trait data on species-level in the future. However, if traits need to be aggregated we showed that trait aggregation is often in conclusion with expert assignments, specifically aggregation approaches using the median. Which aggregation method the objectively better measure is cannot be answered, but when trait aggregation is necessary 1) median aggregation seems to be favourable, 2) highly variable traits and taxa where a few genera have a high number of species should be treated carefully, 3) coding of traits should be considered, and 4) weighting trait information based on the taxa present in the used trait database does not seem to have much influence. No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits, a measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate.
